\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2021-08-01 manuscript.Rnw
Some edits to better compare hesim with related R packages

Despite their computational demands and foundations in statistics, health economic models have historically been developed with specialized commercial software (e.g., \proglang{TreeAge}) or more commonly with a spreadsheet (almost always Microsoft \proglang{Excel}). The limitations of such software relative to programming languages like \proglang{R} have been increasingly emphasized in the literature \citep{baio2017simple, incerti2019r, jalal2017overview}. It is therefore no surprise that a number of related \proglang{R} packages have recently been developed, such as \pkg{BCEA} \citep{baio2017bayesian}, \pkg{SAVI} \citep{strong2014estimating}, \pkg{dampack} \citep{alarid2018dampack}, \pkg{EVSI} \citep{heath2018efficient}, \pkg{survHE} \citep{baio2020survhe}, and \pkg{heemod} \citep{filipovic2017markov}. Still, of the available packages, only \pkg{heemod} provides a general purpose framework for developing simulation models and it is limited to Markov cohort models.

@@

@@

Analyses were performed using \pkg{hesim} version \code{0.5.0}. All code including a replication \proglang{R} script for this paper is available at \url{https://github.com/hesim-dev/hesim-manuscript}.

@@

As mentioned in the introduction, \pkg{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling. It differs from \pkg{hesim} in that it focuses solely on cDTSTMs (i.e., Markov cohort models) and is not designed to facilitate integration of economic models with the statistical models used for parameter estimation. Another important difference is that \pkg{hesim} is considerably faster since all simulations are vectorized (i.e, performed at the \proglang{C++} level). Performance comparisons are provided on the package website. The time-inhomogeneous total hip replacement example from \citet{briggs2006decision} was run with both packages using 1,000 PSA iterations. At the time of writing, a cohort model in \pkg{heemod} completed in approximately 1.9 minutes while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \pkg{hesim} ran in 7.4 seconds.

>>>>

Despite their computational demands and foundations in statistics, health economic models have historically been developed with specialized commercial software (e.g., \proglang{TreeAge}) or more commonly with a spreadsheet (almost always Microsoft \proglang{Excel}). The limitations of such software relative to programming languages like \proglang{R} have been increasingly emphasized in the literature \citep{baio2017simple, incerti2019r, jalal2017overview}. It is therefore no surprise that a number of related \proglang{R} packages have recently been developed, such as \pkg{BCEA} \citep{baio2017bayesian}, \pkg{SAVI} \citep{strong2014estimating}, \pkg{dampack} \citep{alarid2018dampack}, \pkg{EVSI} \citep{heath2018efficient}, \pkg{survHE} \citep{baio2020survhe}, and \pkg{heemod} \citep{filipovic2017markov}.

The functionality of existing packages can generally be separated into three categories: estimation of model parameters, simulation of model outcomes, and decision analysis. Nearly all packages mentioned above support decision analysis in some capacity: \pkg{BCEA}, \pkg{SAVI}, \pkg{dampack}, \pkg{EVSI}, and \pkg{heemod} all contain features for post-processing the output of simulation models in order to quantify decision uncertainty and conduct value of information analyses. On the other hand, there is, to our knowledge, only one package, \pkg{heemod}, that is designed for simulation of health economic models. \pkg{heemod} supports development of Markov cohort models and can incorporate parameter uncertainty for a PSA. Finally, it is important to note that one of, it not the primary advantages of using \proglang{R} for cost-effectiveness modeling is the rich set of packages available for statistical analysis. While any number of these can be used for model parameter estimation---as will be partially demonstrated in this paper---\pkg{survHE} is unique in that it was designed specifically for use in health economic evaluations and consequently contains features such as convenient comparisons of multiple parametric survival models and simulation of model parameters for PSA.

@@

Of the existing packages, \pkg{hesim} is most similar to \pkg{heemod} in that it is designed primarily for simulation of economic models, but can also be used to conduct a CEA. That said, there are some notable differences. First, \pkg{hesim} can be used for individual-level simulations, whereas \pkg{heemod} only supports cohort-level models. Second, all outcomes in \pkg{hesim} are simulated across treatment strategies and patient subgroups within a PSA in a vectorized fashion. This not only increases speed because outcomes are simulated at the \proglang{C++} level, but also makes it easier to account for patient heterogeneity and compare a large number of treatment strategies. Third, since all model types in \pkg{hesim} can be parameterized by fitting statistical models in \proglang{R}, it is easier to integrate the statistical analysis and economic model structure.

@@

Analyses were performed using \pkg{hesim} version \code{0.5.1}. All code including a replication \proglang{R} script for this paper is available at \url{https://github.com/hesim-dev/hesim-manuscript}.

@@

As mentioned in the introduction, \pkg{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling. Unlike \pkg{hesim}, simulations are not vectorized. To asses the significance of this, performance comparisons are provided on the package website. The time-inhomogeneous total hip replacement example from \citet{briggs2006decision} was run with both packages using 1,000 PSA iterations. At the time of writing, a cohort model in \pkg{heemod} completed in approximately 1.9 minutes while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \pkg{hesim} ran in 7.4 seconds.
